### 2013-07-03 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
adding first part of subdomain section

In addition to comparing MCSA scaling to conventional methods, we also
motivate avoiding any type of master/slave algorithms at these high
levels of concurrency by comparing the binary communication tree
implementation used in the previous results with a parallel MCSA
implementation that uses the master/slave scheme to determine the end
of a history transport stage as shown in
Figure~\ref{fig:master_comm_tree}
instead. Figure~\ref{fig:titan_strong_bvsm} provides this comparison
using the strong scaling study while Figure~\ref{fig:titan_weak_bvsm}
gives the results of the comparison for the weak scaling study. In
both cases, the deficiencies caused by using the master/slave scheme
for transport completion are obvious. At around 1,000 cores, both the
strong and weak scaling performance seriously degrades while the
binary tree pattern maintains execellent performance. At this number
of cores, any extra work given to the master process, even if it
amounts to summing a handful of integers and checking for a few extra
messages as in this algorithm, this extra work grows enough to create
serious load imbalance and reduce performance.

@@

Based on the analysis of the pure domain decomposed MCSA algorithm and
subsequently the MSOD implementation, one may ask then ask if there is
any need to commincate histories from domain-to-domain at all. Overlap
was not successful because so few histories needed to be communicated
in the first place that keeping them on-process resulted in more work
than the communications savings that occurred. If that is true, then
perhaps without any overlap or communication, just doing local Monte
Carlo may be satisfactory for accelerating the MCSA iteration.

Full-clip actually uses the solver as a true subdomain solver, just
like you would in a Schwarz method. In fact, this is a stochastic
realization of a Richardson iteration - they are very similar in those
regards.

Overlap is not effective because histories apparently dump most of
their useful information within the first few steps (which is why we
can use such a large weight cutoff and there is an insensitivity to
weight cutoff). So keeping them around to do extra work is not
improving iterative performance - thus increasing compute times. This
is true even for ill-conditioned cases where adding overlap might keep
a number of histories within the domain for significantly longer.

@@

>>>>

In addition to comparing MCSA scaling to conventional methods, we also
motivate avoiding any type of master/slave algorithms at these high
levels of concurrency by comparing the binary communication tree
implementation used in the previous results with a parallel MCSA
implementation that uses the master/slave scheme to determine the end
of a history transport stage as shown in
Figure~\ref{fig:master_comm_tree}
instead. Figure~\ref{fig:titan_strong_bvsm} provides this comparison
using the strong scaling study while Figure~\ref{fig:titan_weak_bvsm}
gives the results of the comparison for the weak scaling study. In
both cases, the deficiencies caused by using the master/slave scheme
for transport completion are obvious. At around 1,000 cores, both the
strong and weak scaling performance seriously degrades while the
binary tree pattern maintains excellent performance. At this number of
cores, any extra work given to the master process, even if it amounts
to summing a handful of integers and checking for a few extra messages
as in this algorithm, grows enough to create serious load imbalance
and reduce performance.

@@

Based on the analysis of the pure domain decomposed MCSA algorithm and
subsequently the MSOD implementation, one may ask then ask if there is
any need to commincate histories from domain-to-domain at all. Overlap
was not successful because so few histories needed to be communicated
in the first place that keeping them on-process resulted in more work
than the communications savings that occurred. If that is true, then
perhaps without any overlap or communication, just doing local Monte
Carlo may be satisfactory for accelerating the MCSA iteration. In this
case, histories near the boundaries of the domain may be truncated
before reaching the weight cutoff if the reach the domain boundary
first. This truncation has the potential to increase the stochastic
uncertainty of the Neumann-Ulam result and therefore increase the
number of MCSA iterations required to converge. In addition, the
scalability of the method will then be purely limited by the parallel
matrix and vector operations required to implement the remainder of
the algorithm.

Considering using MCSA in this way such that the Monte Carlo solve
occurs only in the subdomains casts the method as more of a true
domain decomposition as outlined by Chapter 14 of Saad's text on
iterative methods \citep{saad_iterative_2003}. Compared to traditional
domain decomposition methods we see many similarities including the
potential to overlap the subdomains to improve the results and the
approximate subdomain solves effectively through a relaxation
scheme. By building the boundary data as outlined in
\S~\ref{subsec:domain_generation}, we are effectively gathering one
level of the adjacent domains, albeit in this case to satisfy the data
requirements of the Monte Carlo estimators. Furthermore, the fact that
we are using residual Monte Carlo to compute the correction vector
casts MCSA as a stochastic realization of a additive Schwarz method
with a fixed point iterations used as a relaxation step\footnote{See
  section 14.3.3 in \citep{saad_iterative_2003} for a full definition
  of the additive Schwarz method.}. Both the fixed point iteration and
the computation of the residual propagate the solution from individual
subdomains to the others through the matrix-vector multiply operation
after all subdomain Monte Carlo solves have been completed. Perhaps an
even more important consequence of viewing MCSA as a stochastic
additive Schwarz procedure is that it has the potential to more
readily fit as an accelerator into other linear solver techniques
including GMRES or potentially as a preconditioner.

Using the Neumann-Ulam solver to compute the MCSA correction only in
the subdomains, the strong and weak scaling studies were repeated to
assess the performance gains made by eliminating communication in the
Monte Carlo solver. For each scaling study, the 16-core base case with
Monte Carlo communication was used as the reference point for
efficiency computations in order to measure how eliminating
communication boosts
efficiencies. Figure~\ref{fig:titan_strong_subdomain} gives the
results of the strong scaling study and
Figure~\ref{fig:titan_weak_subdomain} gives the results of the the
weak scaling study in terms of absolute efficiencies. In both cases,
the benefits to parallel performance of using MCSA as a subdomain
solver are obvious. In the strong scaling case we make the important
observation that we hit the strong scaling wall at the same time with
and without the additional Monte Carlo communication. Although the
Monte Carlo communication did contribute to the efficiency decrease,
it is clear that the remaining matrix and vector operations are the
primary impediment to scalability. Furthermore, we see an even greater
rise above unity in efficiencies because there is less communication
overhead hiding the better cache utilization as the local problem
shrinks. At 16,384 cores, performing the subdomain solves boost the
absolute efficiency from 38\% to 67\% relative to the reference case.

In the weak scaling study, we do not have the scaling wall to contend
with and eliminating Monte Carlo communication through subdomain
solves provides an effectively linear boost in the weak scaling
absolute efficiency. We still see a small reduction in weak scaling
efficiency as a function of core count due to the communication
overhead for the parallel matrix and vector operations. Here we
observe super unitary efficiencies for the subdomain case as there was
a substantial improvement in runtimes relative to the 16-core base
case with Monte Carlo communication.

@@

Of additional concern for the subdomain method is the potential
decrease in iterative performance due to the fact that histories near
the boundary will potentially have their random walk truncated, thus
decreasing the information available to compute the correction
vector. For both the weak and strong scaling study, all calculations
converged in 22 or in only a few cases 21 iterations for both the
subdomain solve and the solve with communication giving an algorithmic
efficiency of 1.048 for all cases. This small improvement in iteration
count was not observed to be a function of problem size, number of
processors, or whether the subdomain method was used. Therefore, there
is no reason to expect overlap to improve the iterative or scaling
performance of the method as iterative performance is maintained
without it. Overlap is not effective because histories relinquish most
of their useful information within the first few steps (which is why
we can use such a large weight cutoff and there is an insensitivity to
weight cutoff as shown in
Figure~\ref{fig:estimator_wc_iters}). Keeping them around longer with
overlap to do extra work in the local domain will not improve
iterative performance and thus increasing compute times.

\subsubsection{Strong Scaling with Multiple Sets}
\label{subsubsec:subdomain_ms_strong}

\subsubsection{Weak Scaling with Multiple Sets}
\label{subsubsec:subdomain_ms_weak}
